A Study of Immersion Factors Using an Augmented Reality Simulation

### 2009-06-24 trunk/Paper/paper.tex
outlining
added text about analysis, changed to sections

@@

@@

@@

@@

In our experiment we vary two parameters: the field of view of the AR interface and the length of tracking failure periods.  Each trial lasted sixty seconds, and had seven tracking failures.  The total vertical field of view of our HMD is 36 degrees; the three possible values for the augmented field of view were 10, 20, and 34 degrees.  We varied the length of tracking failures between two seconds, one second and zero seconds (no failure).

@@

\section{Conclusions}

>>>>

Augmented Reality systems may be used to simulate x-ray vision \cite{1383060}.  The ability to see through solid objects has applications in fields ranging from engineering and architecture to military or search-and-rescue operations.  Unfortunately, implementing x-ray vision is not a trivial task.  AR systems have cost issules, including those pertaining to level-of-immersion.  For example, what is the ideal augmented field of view (FOV) for a given scenario?  Other issues relate to tracking the objects in the x-ray view.

motivation??

@@

Livingston and Ai \cite{4637329} studied the effects of various sources of registration error with an AR simulation similar to our own.  Here, participants tracked a virtual car moving throughout a real environment, with a white box representing an augmented view of the car's location.  A white box was continuously visible, even when the car itself was occluded by a building in the environment.  Other virtual cars and their associated white boxes acted as distractors.  At specific times during the experiment, the simulation would freeze and the participant was asked to align the center of their view (indicated by cross-hairs) on the location of the correct car and press a key.  While the effects of several types of error were studied in this work, the effect of registration dropouts was not examined.

Bane and Hollerer \cite{1383060} propose a set of AR tools for virtual x-ray vision.  Various augmented views are presented, each attempting to overcome the "Superman's X-ray vision" problem of presenting too much AR information to the user.  The toolset's usefulness is illustrated with an example of an outdoors user viewing the contents of a nearby building.

other studies which vary AR fov, tracking failures

previous X-ray AR work

simulated AR

@@

\section{AR Simulation}

@@

\begin{figure}[ht!]
	\centering
	\includegraphics[width=3.5in]{annotated_room.png}
	\caption{Overhead view of virtual room (approx. measure).  Participant is stationary in the center of the room, with initial orientation along ``user view'' arrow.  Top-hat's initial position is outside the front doorway, directly in the participant's view.}
\end{figure}

@@

In our experiment we vary two parameters: the field of view of the AR interface and the length of tracking failure periods.  Each trial lasted sixty seconds, and had seven tracking failures.  The total vertical field of view of our HMD is 36 degrees; the three possible values for the augmented field of view were 10, 20, and 34 degrees.  We varied the length of tracking failures between two seconds, one second and zero seconds (no failure).  We experimented with longer values but deemed them too long to be realistic through expert analysis.  This gave us a total of 9 conditions, and for each participant we tested each condition 3 times giving a total of 27 trials per participant.

During each trial there are 20 people in addtition to the man wearing the tophat waking around outside of the room.  The walking path for each person was given by a randomly generated set of points with the constraint that they must stay within the rectangular area of size X by X surrounding the room.  At each point in the path the person would randomly change his walking speed to a new value which was within a range we deemed accecptable through expert analysis.  The man wearing the tophat had the additional constraints as follows.  He must start at the same location for each trial which made him visible by the participant through the door shaped opening in the building.  In order to make the paths by of similar dificulty we decided, through expert analysis, that through the trial he must have walked around the entire building at least once, and he must be visible through the windows for at least $x$ seconds.  We made sure these constraints were met by randomly generating paths until they were satisfied.  We generated 27 sets of patahs this way, one for each trail.  We used a balanced latin square to order these paths and the conditions to discourage the effects of learning.

The experiment was written in python using the Vizard environemnt.  Tracking was done using IntersenseTracker
The HMD we used was a Proview HMD with the following attirbutes. ....  Fill this in.
The computer we used was a Dell with the following specs...........  Fill this in

In our experiment we define performance as effectiveness in tracking the top hat character.  We measure performance by recording the angular distance (in yaw) between the the user's viewing direction and the direction towards the target character.  This measure ranges between zero and 180 degrees (we used the smaller of the two options, clockwise and counter-clockwise).  We record one measurement per video frame, at about sixty frames per second, resulting in about 3600 measurements for a one second trial.

Because the video does not run at exactly 60 Hz, each trial might have a slightly different number of measurements.  We also record a timestamp for each measurement, so that we can determine their exact frequency.  As a pre-processing step, we linearly re-sample the data for each trial so that each has exactly 3600 measurements at 60 Hz.

We observed that participants tend to switch between two different states during the experiment: a tracking state, where they are following the target; and a lost state, where they are searching for the target.  The two states are easily visible in the data.  In the tracking state, the angular error is generally low.  We asked participants to keep the cross hairs on the target as closely as they could, but different participants achieved different levels of accuracy.  In the lost state, the error starts to steadily climb, and may fluctuate depending on where the target moves.  The error may even return to near-zero during a lost state, if the participant unwittingly crosses the target's path.

We hypothesize that less immersive conditions will have longer and more frequent periods where the participant is lost.  We considered several different metrics which might be appropriate in testing this hypothesis.

The simplest is the average error over an entire trial.  However, this may not accurately capture the amount that a participant is lost, because the error does not necessarily stay high during the lost state, and also different participants may generally keep the cross hair further from the target even in the tracking state.

We can more explicitly detect the lost state by applying a threshold $\tau$ to the data.  When the error rises above the threshold, we consider the participant lost.  TODO: how to determine this threshold??

One metric which uses this threshold is the ``time to failure,'' the time until a lost state is encountered.  However a participant may reach the lost state sooner or later depending on the movement of the avatars.  We also considered a measurement we called ``time tracking,'' which is the amount of time spent in the tracking state.  This metric seems to most clearly represent how well a participant performed.  Trials with either more lost periods or longer lost periods will result in a lower total time tracking.

@@

\section{Conclusions and Future Work}
